\documentclass[../main.tex]{subfiles}

\begin{document}

\chapter{Particle Interaction Physics}
The mathematical models derived in the previous chapters form the basis of our light ion transport model. In this and the following chapters the interaction physics and multigroup data describing the interactions of light ions as they slow down are discussed. This data will then be combined with our mathematical models to solve light ion transport equations in ICF applications.

A general collision between an incident energetic light ion of species $x$ and a target charged particle of species $T$ resulting in species $s$ and $r$ is written as
\begin{equation} \label{eqn:general-collision}
    x+t \rightarrow r+s \equiv T(x,s)R,
\end{equation}
where particles $s$ and $R$ are termed the scattered and recoiling particles. The general differential cross section that describes the collision in Eq. \eqref{eqn:general-collision} is
\begin{equation} \label{eqn:differentialCrossSection}
    \sigma_{x,t \rightarrow s}(E^{\prime} \rightarrow E; \mucm),
\end{equation}
where the incident ion, $x$, had energy $E^{\prime}$ in the laboratory (LAB) frame of reference before its collision with the target species $t$. The resulting species $s$ is scattered through an angle whose center of mass (CM) cosine is $\mucm$ and has resulting energy $E$ in the LAB frame of reference. The differential cross section in Eq. \eqref{eqn:differential cross section} is a microscopic differential cross section and has units of $cm^{2}$, and is related to the macroscopic differential cross section as
\begin{equation}
    \Sigma_{x,t \rightarrow s}(E^{\prime} \rightarrow E; \mucm) = \mathcal{N} \sigma_{x,t \rightarrow s}(E^{\prime} \rightarrow E; \mucm),
\end{equation}
where $\mathcal{N}$ is the number density of the target species in $cm^{-3}$.

In Eqs. \eqref{eqn:general-collision} and \eqref{eqn:differentialCrossSection}, collisions where the resulting species is identical to the incident species, $s = x$, are termed elastic scattering collisions; and collisions where the resulting species differs from the incident species, $s \neq x$, are termed nuclear-reaction collisions. These nuclear-reaction collisions result in the destruction of the incident ions and are therefore considered absorption collisions for the incident ion. Furthermore, ion collisions can either be described as light or heavy depending on whether the incident ion is lighter or heavier than the target particle respectively. The heavy differential cross sections are related to the light differential cross sections through kinematics and are discussed in the next chapter, as such we only discuss light collisions in this chapter.

As energetic light ions slow down in they experience three different types of collisions with the thermal background ions and electrons: 
\begin{enumerate}
    \item elastic collisions with thermal electrons;
    \item elastic collisions with thermal ions; or
    \item thermonuclear reaction collisions with thermal ions.
\end{enumerate}

% ------------------------------------------------
% ION ELASTIC SCATTERING
% ------------------------------------------------
\input{chapters/chapter-4/ionElasticScattering.tex}

% ------------------------------------------------
% NUCLEAR REACTION DIFFERENTIAL CROSS SECTIONS 
% ------------------------------------------------
\input{chapters/chapter-4/nuclearReaction.tex}

% ------------------------------------------------
% NDI STOPPING POWERS
% ------------------------------------------------
\input{chapters/chapter-4/ndiStoppingPowers.tex}

% ------------------------------------------------
% NEUTRON ELASTIC SCATTERING
% ------------------------------------------------
\input{chapters/chapter-4/incidentNeutron.tex}

\end{document}
