\documentclass[../main.tex]{subfiles}

\begin{document}

\chapter{Charged Particle Interaction Physics}
The interaction physics that describe the slowing down of energetic light ions through collisions with background charged particles in plasmas are described in terms of differential cross sections. A general collision between an incident energetic light ion of species $x$ and a target charged particle of species $t$ resulting in new species $s$ and $r$ is written as
\begin{equation} \label{eqn:general-collision}
    t(x,s)r,
\end{equation}
where particles $s$ and $r$ are termed the scattered and recoiling particles. The general differential cross section that describes the collision in Eq. \eqref{eqn:general-collision} is
\begin{equation} \label{eqn:differential cross section}
    \sigma_{x,t \rightarrow s}(E^{\prime} \rightarrow E; \mucm),
\end{equation}
where the incident ion, $x$, had energy $E^{\prime}$ before its collision with the target species $t$. The resulting species $s$ is scattered through an angle whose center of mass (CM) cosine is $\mucm$ and has resulting energy $E$. The differential cross section in Eq. \eqref{eqn:differential cross section} is a microscopic differential cross section and has units of $cm^{2}$, and is related to the macroscopic differential cross section as
\begin{equation}
    \Sigma_{x,t \rightarrow s}(E^{\prime} \rightarrow E; \mucm) = \mathcal{N} \sigma_{x,t \rightarrow s}(E^{\prime} \rightarrow E; \mucm),
\end{equation}
where $\mathcal{N}$ is the number density of the target species in $cm^{-3}$.

Collisions where the resulting species is identical to the incident species, $s = x$, are termed elastic scattering collisions; and collisions where the resulting species differs from the incident species, $s \neq x$, are termed nuclear-reaction collisions. These nuclear-reaction collisions result in the destruction of the incident ions and are therefore considered absorption collisions for the incident ion. Furthermore, ion collisions can either be described as light or heavy depending on whether the incident ion is lighter or heavier than the target particle respectively. 

As energetic ions slow down they experience four types of collisions with the background charged particles present in the plasma:
\begin{enumerate}
    \item elastic-Coulomb collisions with thermal electrons;
    \item elastic-Coulomb collisions with thermal ions;
    \item nuclear-elastic scattering (NES) collisions with thermal ions; or
    \item nuclear-reaction collisions with thermal ions.
\end{enumerate}
Because the ion-ion elastic scattering collisions consist of two parts (Coulomb and NES), a quantum mechanical interference differential cross section appears that describes the mutual interference between the two types of elastic scattering collisions \cite{Devaney-1971}. Therefore, the total elastic scattering differential cross section is written as
\begin{equation}
    \sigma_{x,t \rightarrow x} = \sigma_{x,t \rightarrow x}^C + \sigma_{x,t \rightarrow x}^N + \sigma_{x,t \rightarrow x}^I,
\end{equation}
where the superscripts $C$, $N$, and $I$ represent the Coulomb, NES, and interference components of the total ion-ion elastic scattering differential cross section. The differential cross sections that describe Coulomb collisions have an analytical form; while the differential cross sections that describe the NES, interference, and nuclear reactions collisions are tabulated in data sets. The remainder of this section presents the analytical form of the Coulomb differential cross section, and the forms of the tabulated differential cross sections describing the NES, interference, and nuclear-reaction collisions.

% ------------------------------------------------
% COULOMB DIFFERENTIAL CROSS SECTIONS
% ------------------------------------------------
\input{interactionPhysics/coulombDCS.tex}

% ------------------------------------------------
% NES DIFFERENTIAL CROSS SECTIONS 
% ------------------------------------------------
\input{interactionPhysics/nuclearElasticScattering.tex}

% ------------------------------------------------
% INTERFERENCE DIFFERENTIAL CROSS SECTIONS 
% ------------------------------------------------
\input{interactionPhysics/interference.tex}

% ------------------------------------------------
% NUCLEAR REACTION DIFFERENTIAL CROSS SECTIONS 
% ------------------------------------------------
\input{interactionPhysics/nuclearReaction.tex}

\end{document}
